On consistency scores in text data with an implementation in R

### 2020-09-03 outputs/documents/paper_ngrams.Rmd
paper minor updates

We employ n-gram models to calculate consistency scores and generate word candidates for text corrections. The n-gram approach involves identifying words that are commonly found together. In our application, we use these sequences to calculate consistency scores and generate word candidates for text corrections. For instance, if using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset. To calculate consistency scores and generate word candidates for text corrections, we use mixed sequences of five, four and three consecutive words also known as 5-grams, 4-grams and tri-grams.

@@

In this paper, we introduce consistency scores and an R package that implements them. In Section 2, we provide an overview of language models and n-grams in particular. In Section 3 we deconstruct the package functions and provide demonstration example of the package usage. Finally, in Section 4 we discuss limitations and future directions.

@@

Given the nature of human languages, some combinations of words tend to occur more frequently than others. Think of 'good', which is more often followed by 'morning', than 'duck'. As such, we could consider English text production as a conditional probability, $\mbox{Pr}(w_{k} | w^{k-1}_{1})$, where $k$ is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence [@brown1992class]. In this way, the generation of some prediction, $w_{k}$, is based on the history, $w^{k-1}_{1}$. This is the underlying principle of all language models. Essentially, a language model is a probability distribution over sequences of words. The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents [@bengio2003neural]. However, this is difficult as language is categorical. If we consider each word in a vocabulary as a category, then the dimensionality of language becomes large [@rosenfeld2000two]. The reason that there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem.

@@

In our application, we use a model that has a mixture of 5-grams, 4-grams and tri-grams to provide word replacements to a text error. Longer n-grams such as 5-grams would capture more context of the sequence and potentially generate more accurate predictions. However, there is less probability for a longer sequence of words to fit a specific combination. Therefore, shorter n-grams should still be included to preserve generalizability of the predictions.

@@

In the 4-gram model, one extra more word "elephant" is included and thus provides more context for the prediction, and the word "room" is predicted by the 4-gram model. However, there are times people would simply describe the location of an elephant, for example "elephant in the pond". This combination of this sequence as a 4-gram might not be in a comparison dataset. Therefore, we have to also rely on the tri-gram models to provide other word candidates including "in the pond".

@@

To obtain the internal consistency score, we will construct a internal consistency dataset from a body of text that is the source of the text being evaluated. The function `make_internal_consistency_dataset` is created for this purpose. The function turns the body_of_text into an internal consistency dataset which the text being evaluated is compared with. To do so, the function first transforms the body of text into word tokens with the punctuation removed and the letters turned to lowercase:

@@

Finally, the `make_internal_consistency_dataset` function splits the n-grams to two parts: the previous $n-1$ words and the last word. In the end, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:

@@

After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function. Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into 5-grams, 4-grams and tri-grams tokens, then splits the n-grams and preserves them into a tibble:

@@

Next, the function combines the text tokens tibble and the internal consistency tibble by the `left_join()` function of `dplyr`. The `left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles. By combining the two tibbles, the function generates two additional columns — the `last_word` column are the last words of all tri-grams, and the `last_word_expected` contains the words that are in the internal consistency dataset. By comparing the words in `last_word` and `last_word_expected`, we can identify the "unexpected" words. If a last word is in `last_word` but not `last_word_expected`, it is regarded as an unexpected word and the corresponding word in `last_word_expected` serves as a replacement candidate:

@@

The `generate_external_consistency_score` function works identically with the `generate_internal_consistency_score`. The only difference is that it compares the text being evaluated with an external consistency dataset that is larger and more general. The external consistency dataset we employed in the application is constructed using a subset text data from Open WebText Corpus by @Gokaslan2019OpenWeb.

@@

The first step is to turn the body of text into the internal consistency dataset. The generated internal consistency dataset is a tibble that contains the 5-grams, 4-grams and tri-grams that appear in the internal text data more than once. Since only `there_was_no` has more than one occurrence, it is the only reference tri-gram in the internal consistency dataset:

@@

To get the external consistency score, we will employ the `generate_external_consistency_score` function. The function `generate_internal_consistency_score` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset. Here, the text to evaluate is "there was no possibiliti", with possibility being wrongly spelled. The function identifies the word "possibiliti" as an unexpected word, and generates the external consistency score as 0.833. Because the external dataset is larger, there are several words generated as the replacement candidates of "possibiliti". The replacements based on 4-grams are displayed first followed by results based on tri-grams:

@@

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class]. As demonstrated in our application, we use a tri-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates. However, n-gram models do not take the linguistic structure of language into account. For instance, @rosenfeld2000two [p. 1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. The prediction of the next word is based on only a few preceding words and broader context is not taken into account. Hence next-word prediction using n-gram based language models can be limited. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance, an equally reasonable two-gram involving the word 'good' is 'good work'. However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morning' or 'work' is to be suggested as the next word after 'good'.

@@

We also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language] and GPT-3 [@brown2020language] and BERT [@devlin2018bert]. GPT-2, GPT-3 and BERT are pre-trained language models that are built based on the Transformer architecture [@vaswani2017attention]. Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention]. Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing]. Therefore, we can say that Transformer marks a new era of language models. In our future works, it will be important for us to implement different language models compare their effectiveness in the tasks of generating consistency score and providing suggested text corrections and improve the `aRianna` package.

>>>>

We employ n-gram models to calculate consistency scores and generate word candidates for text corrections. The n-gram approach involves identifying words that are commonly found together. In our application, we use these sequences to calculate consistency scores and generate word candidates for text corrections. For instance, if using a sequence of three words, or tri-grams, then we identify the first two words in the sequence and evaluate if the last word in the sequence is 'as expected' by comparing it to the corresponding tri-gram in a comparison dataset. The sequences we employed are a mixture of five, four and three consecutive words also known as 5-grams, 4-grams and tri-grams.

@@

In this paper, we introduce consistency scores and an R package that implements them. In Section 2, we provide an overview of language models and n-grams in particular. In Section 3, we deconstruct the package functions and provide demonstration example of the package usage. Finally, in Section 4, we discuss limitations and future directions.

@@

Given the nature of human languages, some combinations of words tend to occur more frequently than others. Think of 'good', which is more often followed by 'morning', than 'duck'. As such, we could consider English text production as a conditional probability, $\mbox{Pr}(w_{k} | w^{k-1}_{1})$, where $k$ is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence [@brown1992class]. In this way, the generation of some prediction, $w_{k}$, is based on the history, $w^{k-1}_{1}$. This is the underlying principle of all language models. Essentially, a language model is a probability distribution over sequences of words. The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents [@bengio2003neural]. However, this is difficult as language is categorical. If we consider each word in a vocabulary as a category, then the dimensionality of language becomes large [@rosenfeld2000two]. The reason that there is such a variety of statistical language models is that there are various ways of dealing with this fundamental problem. N-gram is one of the statistical language models that works by taking the co-occurrence of words in sequence into account.

@@

In our application, we use a model that has a mixture of 5-grams, 4-grams and tri-grams to provide word replacements to a text error. Longer n-grams such as 5-grams capture more context of word sequences and potentially generate more accurate predictions. However, there is less probability for a longer sequence of words to fit a specific combination and leads to data sparsity. Therefore, shorter n-grams should still be included to preserve generalizability of the predictions.

@@

In the 4-gram model, one extra more word "elephant" is included and thus provides more context for the word "room" to be accurately predicted. However, there are times people would simply describe the location of an elephant; for example, "elephant in the pond". The combination of this sequence as a 4-gram token might not be in a comparison dataset and no prediction would be available if only the 4-gram models are used. Therefore, we have to also rely on the tri-gram models to provide other word candidates including "in the pond".

@@

To obtain the internal consistency score, we will construct an internal consistency dataset from a body of text that is the source of the text being evaluated. The function `make_internal_consistency_dataset` is created for this purpose. The function turns the body_of_text into an internal consistency dataset which the text being evaluated will be compared with. To do so, the function first transforms the body of text into word tokens with the punctuation removed and the letters turned to lowercase:

@@

Finally, the `make_internal_consistency_dataset` function splits the n-grams to two parts: the previous $n-1$ words and the last word. As a result, a tibble is created with each n-gram containing its original sequence, the first words, and the last word:

@@

After we have the internal consistency dataset, we need a function to compare the text being evaluated and the internal dataset in order to retrieve the internal consistency score — the `generate_internal_consistency_score` function. Similar to the `make_internal_consistency_dataset` function, the function first transforms the text being evaluated into 5-grams, 4-grams and tri-grams tokens, then splits the n-grams to their first words and last words:

@@

Next, the function combines the text tokens tibble and the internal consistency tibble by the `left_join()` function of `dplyr`. The `left_join()` function returns all the rows from the text tokens tibble and all the columns from both tibbles [@2018dplyr]. By combining the two tibbles, the function generates two additional columns — the `last_word` column are the last words of all n-grams, and the `last_word_expected` contains the last words that are in the internal consistency dataset. By comparing the words in `last_word` and `last_word_expected`, we can identify the "unexpected" words. If a last word is in `last_word` but not `last_word_expected`, it is regarded as an unexpected word and the corresponding word in `last_word_expected` serves as a replacement candidate:

@@

The `generate_external_consistency_score` function works identically as the `generate_internal_consistency_score` does. The only difference lies in that it compares the text being evaluated with an external consistency dataset that is larger and more general. The external consistency dataset we employed in the application is constructed using a subset text data from Open WebText Corpus by @Gokaslan2019OpenWeb.

@@

The first step is to turn the body of text into the internal consistency dataset. The generated internal consistency dataset is a tibble that contains the 5-grams, 4-grams and tri-grams that appear in the internal text data more than once. Since only `there_was_no` has more than one occurrence, it is the only reference n-gram in the internal consistency dataset:

@@

To get the external consistency score, we will employ the `generate_external_consistency_score` function. The function `generate_internal_consistency_score` takes in only one argument — the text to evaluate, and compares it with the external consistency dataset. Here, the text to evaluate is "there was no possibiliti", with the word "possibility" being wrongly spelled. The function identifies the word "possibiliti" as an unexpected word, and generates the external consistency score 0.833. Because the external dataset is larger, it is capable of providing more replacement candidates for "possibiliti". The replacements based on 4-grams are displayed first followed by the replacements based on tri-grams:

@@

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class]. As demonstrated in our application, we use a n-gram based model to calculate internal and external consistency scores and generate a list of words as text correction candidates. However, n-gram models do not take the linguistic structure of language into account. For instance, @rosenfeld2000two [p. 1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. The prediction of the next word is based on only a few preceding words and broader context is not taken into account. Hence next-word prediction using n-gram based language models can be limited. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance, an equally reasonable two-gram involving the word 'good' is 'good work'. However, because n-gram models does not take the surrounding text into account, it has not capacity of judging if 'morning' or 'work' is to be suggested as the next word after 'good'.

@@

We also consider incorporating generative pre-trained transformer models such as GPT-2 [@radford2019language], GPT-3 [@brown2020language], and BERT [@devlin2018bert]. GPT-2, GPT-3 and BERT are pre-trained language models that are built based on the Transformer architecture [@vaswani2017attention]. Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention]. Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing]. Therefore, we can say that Transformer marks a new era of language models. In our future works, it will be important for us to implement different language models compare their effectiveness in the tasks of generating consistency score and providing suggested text corrections. We intend to use our future findings to continuously improve the `aRianna` package.
